---
title: "Lab 3 Report"
author: "Tyler Edvall"
date: "`r format(Sys.Date())`"
output: github_document
---

Instructions for this lab report can be found on [Your assignment section of Lab Assignment 3 Introduction to ggplot2](https://biol275-msum.github.io/introduction-to-ggplot2.html#your-assignment) on the lab website.

## Fireflies

> A. Insert an R code chunk and create a graph depicting the frequency distribution of the 35 mass measurements. It should have legible text and appropriate axis labels.

```{r}

# spermatophore mass --------------------------------------------------------

firefly_data <- read_csv("https://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q19FireflySpermatophoreMass.csv")

firefly_data                  

ggplot(data = firefly_data)

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass))

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.02)

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.02,
                 boundary = 0, closed = "left")

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.02,
                 boundary = 0, closed = "left")

ggplot(data = firefly_data) +
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.02,
                 boundary = 0, closed = "left", 
                 fill = "#C5351B", color = "black") +
  labs(x = "Spermatophore Mass (Mg)", y = "Frequency (number of fireflies)") +
  
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1))
  )
  
```

> B. What type of graph did you choose in part (A)? Why?

We chose a histogram to plot the firefly data. The reason is because the firefly data is numerical and a histogram is the best way to show this type of data. 

> C. Describe the shape of the frequency distribution. What are its main features?

The shape of the frequency distribution is skewed to the left. There is a noticeable outlier to the right of the graph. 

> D. What term would be used to describe the largest measurement in the frequency distribution? You can refer to the lecture slides, your notes, or the textbook.

The mode. 

## Bird orders

I read the data using the code provided by the instructor. The dataset is from the auk package and was originally named ebird_taxonomy. I converted it to a tibble, removed non-species taxa, and named the new dataset birds.


library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa

> E. How many bird species are in the new birds dataset? How did you find out?

There are 10721 bird species. I figured this out by looking under the observation tab and seeing the number of observations.

> H. How many orders are there? You can get this right by counting on the graph, but you will only get full points if you use some code to figure it out.

41 different orders.

```{r}
nrow(distinct(birds, order))
```

> F. Insert an R code chunk and create a graph depicting the distribution of orders in the birds dataset. Sort the orders with the most frequent on the left. It should have legible text and appropriate axis labels.

```{r}

# Bird orders data ------------------------------------------------------

library(auk)                          # load the auk package
birds <- ebird_taxonomy %>%           # start with the ebird_taxonomy data
  as_tibble() %>%                     # tibbles print better in the console
  filter(category == "species")       # remove non-species taxa

birds

distinct(birds, order)

count(birds, family_common)

ggplot(data = birds)

ggplot(data = birds) +
  geom_bar(mapping = aes(x = order))

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)))

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B")

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B") +
  labs(x = "Order", y = "Frequency (number of birds)")

ggplot(data = birds) +
  geom_bar(mapping = aes(x = fct_infreq(order)), fill = "#C5351B", 
           width = .8) +
  labs(x = "Order", y = "Frequency (number of birds)") +
  scale_y_continuous(limits = c(0, 7000), expand = expansion(mult = 0)) +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1)),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_blank()
  )

```

> G. What kind of graph did you create? Why?

Bar graph, since the bird data was categorical. That would be the best way to display the data.

## Links to peer review

I reviewed Kelsey Leach's lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-kelseyaleach

Kelsey Leach reviewed my lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-tdedvall.git

## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```
